### 2023-11-27 src/preprocess.Rmd
Add age ranges, await fix for neo age range data

---
title: "Preprocess"
date: "`r format(Sys.time(), '%B %d, %Y')`"
author: "Richard Wen <richard.wen@unityhealth.to>"
output: html_notebook
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
  message=FALSE,
  warning=FALSE
)
options(readr.show_col_types = FALSE)
```

```{r, echo=FALSE, eval=FALSE}
install.packages("tidyverse")
```

## Libraries

```{r}
library(tidyverse)
```

## Data

Load data from `data` and `tmp` folders.

```{r}
# Physicians cols to keep and rename for consistency
phy_cols <- c(
    "rowid",
    "death_age_group",
    "p1_icd",
    "p2_icd",
    "p1_recon_icd",
    "p2_recon_icd",
    "adj_icd"
)
phy_rename <- c(
    age_range = "death_age_group"
)
phy_adult_cols <- c(
    "rowid",
    "death_age_group",
    "p1_icd_cod",
    "p2_icd_cod",
    "p1_recon_icd_cod",
    "p2_recon_icd_cod",
    "adj_icd_cod"
)
phy_adult_rename <- c(
    age_range = "death_age_group",
    p1_icd = "p1_icd_cod",
    p2_icd = "p2_icd_cod",
    p1_recon_icd = "p1_recon_icd_cod",
    p2_recon_icd = "p2_recon_icd_cod",
    adj_icd = "adj_icd_cod"
)

# Read data
df <- list(
    wva2016_icd10 = read_csv(
        "../data/wva2016_icd10_v1.csv",
        col_types = cols(wva2016_code = col_double())
    ),
    icd10_cghr10 = read_csv("../data/icd10_cghr10_v1.csv"),
    physician = bind_rows(
        read_csv(
            "../tmp/healsl_rd1_adult_v1.csv",
            col_select = all_of(phy_adult_cols)
        ) %>%
            mutate(round = 1, age = "adult") %>%
            rename(all_of(phy_adult_rename)),
        read_csv(
            "../tmp/healsl_rd2_adult_v1.csv",
            col_select = all_of(phy_adult_cols)
        ) %>%
            mutate(round = 2, age = "adult") %>%
            rename(all_of(phy_adult_rename)),
        read_csv(
            "../tmp/healsl_rd1_child_v1.csv",
            col_select = all_of(phy_cols)
        ) %>%
            mutate(round = 1, age = "child") %>%
            rename(all_of(phy_rename)),
        read_csv(
            "../tmp/healsl_rd2_child_v1.csv",
            col_select = all_of(phy_cols)
        ) %>%
            mutate(round = 2, age = "child") %>%
            rename(all_of(phy_rename)),
        read_csv(
            "../tmp/healsl_rd1_neo_v1.csv",
            col_select = all_of(phy_cols)
        ) %>%
            mutate(round = 1, age = "neo") %>%
            rename(all_of(phy_rename)),
        read_csv(
            "../tmp/healsl_rd2_neo_v1.csv",
            col_select = all_of(phy_cols)
        ) %>%
            mutate(round = 2, age = "neo") %>%
            rename(all_of(phy_rename))
    ),
    model_chatgpt3 = bind_rows(
        read_csv(
            "../tmp/healsl_rd1_rapid_chatgpt3_v1.csv",
            col_select = c("rowid", "round", "api_chat_icd")
        ),
        read_csv(
            "../tmp/healsl_rd2_rapid_chatgpt3_v1.csv",
            col_select = c("rowid", "round", "api_chat_icd")
        )
    ) %>% rename(chatgpt3_icd10 = api_chat_icd),
    model_chatgpt4 = bind_rows(
        read_csv(
            "../tmp/healsl_rd1_rapid_chatgpt4_v1.csv",
            col_select = c("rowid", "round", "api_chat_icd")
        ),
        read_csv(
            "../tmp/healsl_rd2_rapid_chatgpt4_v1.csv",
            col_select = c("rowid", "round", "api_chat_icd")
        )
    ) %>% rename(chatgpt4_icd10 = api_chat_icd),
    model_insilicova = bind_rows(
        read_csv(
            "../tmp/healsl_rd1_rapid_insilicova_v1.csv",
            col_select = c("rowid", "round", "cause1_wva")
        ),
        read_csv(
            "../tmp/healsl_rd2_rapid_insilicova_v1.csv",
            col_select = c("rowid", "round", "cause1_wva")
        )
    ) %>% rename(insilicova_wva2016 = cause1_wva),
    model_interva5 = bind_rows(
        read_csv(
            "../tmp/healsl_rd1_rapid_interva5_v1.csv",
            col_select = c("rowid", "round", "cause1_wva")
        ),
        read_csv(
            "../tmp/healsl_rd2_rapid_interva5_v1.csv",
            col_select = c("rowid", "round", "cause1_wva")
        )
    ) %>% rename(interva5_wva2016 = cause1_wva)
)
```

```{r, echo=FALSE}
for (i in names(df)) {
    print(i)
    print(head(df[[i]]))
}
```

## Methods

### Final ICD

Create one column for final physician ICD-10 codes and one column each for physician agreed, reconciled, and adjudicated records based on the following rule:

1. If an adjudicated code exists, use it as the final code
2. If not, use physician 1's reconciled code if it exists
3. Otherwise use physician 1's code

```{r}
df$physician <- df$physician %>%
    mutate(
        physician_icd10 = coalesce(
            adj_icd,
            p1_recon_icd,
            p1_icd
        ),
        is_agreed = if_else(
            is.na(p1_recon_icd) & is.na(p2_recon_icd) & is.na(adj_icd),
            TRUE,
            FALSE
        ),
        is_recon = if_else(
            !is.na(p1_recon_icd) | !is.na(p2_recon_icd),
            TRUE,
            FALSE
        ),
        is_adj = if_else(
            !is.na(adj_icd),
            TRUE,
            FALSE
        )
    ) %>%
    select(
        rowid,
        round,
        age,
        p1_icd,
        p2_icd,
        p1_recon_icd,
        p2_recon_icd,
        adj_icd,
        is_agreed,
        is_recon,
        is_adj,
        physician_icd10
    )
head(df$physician)
```

### Combine Outputs

Combine all model and physician predictions with their ages.

```{r}
models <- df[str_starts(names(df), "model_")]
out <- df$physician
for (i in 1:length(models)) {
    out <- out %>%
        left_join(
            models[[i]],
            by = c("rowid" = "rowid", "round" = "round"),
            suffix = c("", "")
        )
}
head(out)
```

### Convert to CGHR-10

Convert ICD-10 and WVA-2016 codes to CGHR-10 codes.

```{r}

# Get wva2016 cols for model preds
wva_cols <- colnames(out)
wva_cols <- wva_cols[!wva_cols %in% c("rowid", "round", "age")]
wva_cols <- wva_cols[str_ends(wva_cols, "_wva2016")]

# Convert wva2016 to icd10 via join
for (wcol in wva_cols) {

    # Create new converted col based on model name
    mname <- str_remove(wcol, "_wva2016")
    iname <- paste0(mname, "_icd10")

    # Join model wva to mapping wva
    out <- out %>%
    left_join(
        df$wva2016_icd10 %>%
            select(wva2016_code, icd10_code) %>%
            rename_with(~iname, icd10_code),
        by = setNames("wva2016_code", wcol),
        multiple = "first"
    )
}

# Get icd10 cols for model preds
icd_cols <- colnames(out)
icd_cols <- icd_cols[!icd_cols %in% c("rowid", "round", "age")]
icd_cols <- icd_cols[str_ends(icd_cols, "_icd10")]

# Convert icd10 to cghr10 via join
for (icol in icd_cols) {

    # Create new converted col based on model name
    mname <- str_remove(icol, "_icd10")
    cname <- paste0(mname, "_cghr10")

    # Remove all whitespaces and keep only first icd10 code if multiple
    out[[icol]] <- str_replace_all(out[[icol]], "\\s", "")
    out[[icol]] <- str_extract(out[[icol]], "^[A-Za-z]\\d{2}")

    # Join model wva to mapping wva
    out <- out %>%
    left_join(
        df$icd10_cghr10 %>%
            select(icd10_code, cghr10_title, cghr10_age) %>%
            rename_with(~cname, cghr10_title),
        by = setNames(c("icd10_code", "cghr10_age"), c(icol, "age")),
        multiple = "first"
    )
}

head(out)
```

## Results

Save pre-processed data to `data` folder.

```{r}
write_csv(out, "../data/healsl_rd1to2_cod_v1.csv")
```