### 2023-12-01 src/metrics.Rmd
Add cod plot and fix k for pccc

---
title: "Metrics"
date: "`r format(Sys.time(), '%B %d, %Y')`"
author: "Richard Wen <richard.wen@unityhealth.to>"
output: html_notebook
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
  message=FALSE,
  warning=FALSE
)
options(readr.show_col_types = FALSE)
```

```{r, echo=FALSE, eval=FALSE}
install.packages("tidyverse")
```

## Libraries

```{r}
library(tidyverse)
```

## Settings

```{r}
models_format <- c(
    "_cghr10" = "",
    "_" = " & ",
    "chatgpt3" = "ChatGPT-3.5",
    "chatgpt4" = "ChatGPT-4",
    "insilicova" = "InSilicoVA",
    "interva5" = "InterVA-5"
)
stage_format <- c(
    "is_" = "",
    "recon" = "Reconciliation",
    "agreed" = "Agreement",
    "adj" = "Adjudication"
)
```

## Functions

Functions for PCCC and CSMF Accuracy.

```{r}

# Calc PCCC
calc_pccc <- function(actual, pred, k = 1, N = NULL) {

    # Calc N num of causes if not known
    N <- if (is.null(N)) length(unique(na.omit(c(actual, pred)))) else N

    # Calc frac of deaths in top k causes
    TP <- actual == pred
    TP[is.na(TP) | is.null(TP)] <- FALSE # for no preds
    C <- sum(TP) / length(actual)

    # Calc PCCC
    out <- (C - (k/N)) / (1 - (k/N))
    return(out)
}

# Calc CSMF accuracy
calc_csmf_acc <- function(actual, pred) {

    # Get all unique causes
    causes <- unique(c(actual, pred))

    # Get csmfs
    cases <- length(actual)
    csmf_true <- table(actual) / cases
    csmf_pred <- table(pred) / cases

    # Correct for missing causes in either actual or pred
    csmf_true <- vapply(
        causes,
        function(x) if (x %in% names(csmf_true)) csmf_true[x] else 0,
        FUN.VALUE = numeric(1)
    )
    csmf_pred <- vapply(
        causes,
        function(x) if (x %in% names(csmf_pred)) csmf_pred[x] else 0,
        FUN.VALUE = numeric(1)
    )

    # Calc csmf max error
    csmf_max_error <- 2 * (1 - min(csmf_true))

    # Calc csmf acc
    out <- 1 - (sum(abs(csmf_true - csmf_pred)) / csmf_max_error)
    return(out)
}

# Bulk calc per model
calc_per_model <- function(actual, pred, func, name = "Metric", ...) {

    # Prep data in long format grouped by model
    out <- pred %>%
        pivot_longer( # to long format
            everything(),
            names_to = "Model",
            values_to = "prediction"
        ) %>% group_by(Model)

    # Calc k if pccc as it is diff for model combos
    if (identical(func, calc_pccc)) {
        out <- out %>%
            summarise(
                !!name := func(
                    actual,
                    prediction,
                    k = str_count(unique(Model), "_"),
                    ...
                )
            )
    } else { # otherwise apply func
        out <- out %>%
            summarise(
                !!name := func(
                    actual,
                    prediction,
                    ...
                )
            )
    }

    # Format display and return
    out <- out %>%
        mutate(Model = str_replace_all(Model, models_format)) %>%
        arrange(across({{name}}, desc))
    return(out)
}

# Bulk calc per model based on age, physician coding stage, and cause of death
calc_by <- function(
        df,
        func,
        name = "Metric",
        by_age = TRUE,
        by_stage = TRUE,
        by_cod = TRUE,
        ...
    ) {
    out <- list()

    # Calc overall metric
    out[[name]] <- calc_per_model(
        actual = df %>% pull(physician_cghr10),
        pred = df %>% select(ends_with("_cghr10"), -physician_cghr10),
        func = func,
        ...
    ) %>% rename(
        {{ name }} := Metric
    )

    # Calc metric for va coding stage
    if (by_stage) {
        for (stage in c("is_agreed", "is_recon", "is_adj")) {

            # Filter data for stage
            df_filter <- df %>% filter(.[[stage]] == TRUE)
            metric_name = paste0(
                name,
                " ",
                str_replace_all(stage, stage_format)
            )

            # Calc metric for stage
            out[[metric_name]] <- calc_per_model(
                actual = df_filter %>%
                    pull(physician_cghr10),
                pred = df_filter %>%
                    select(ends_with("_cghr10"), -physician_cghr10),
                func = func,
                ...
            ) %>% rename(
                 {{ metric_name }} := Metric
            )
        }
    }

    # Calc pcc for each age group
    if (by_age) {
        for (age_group in c("adult", "child", "neo")) {

            # Filter data for age
            df_filter <- df %>% filter(age == age_group)
            metric_name = paste0(
                name,
                " ",
                str_to_title(age_group)
            )

            # Calc metric for age
            out[[metric_name]] <- calc_per_model(
                actual = df_filter %>%
                    pull(physician_cghr10),
                pred = df_filter %>%
                    select(ends_with("_cghr10"), -physician_cghr10),
                func = func,
                ...
            ) %>% rename(
                 {{ metric_name }} := Metric
            )

            # Calc metric for each stage within age group
            for (stage in c("is_agreed", "is_recon", "is_adj")) {

                # Filter data for stage within age group
                df_filter <- df %>% filter(
                    .[[stage]] == TRUE & age == age_group
                )
                metric_name = paste0(
                    name,
                    " ",
                    str_to_title(age_group),
                    " ",
                    str_replace_all(stage, stage_format)
                )

                # Calc metric for stage within age group
                out[[metric_name]] <- calc_per_model(
                    actual = df_filter %>%
                        pull(physician_cghr10),
                    pred = df_filter %>%
                        select(ends_with("_cghr10"), -physician_cghr10),
                    func = func,
                    ...
                ) %>% rename(
                     {{ metric_name }} := Metric
                )
            }
        }
    }

    # Calc metric by cause of death
    if (by_cod) {

        # Get unique causes of death
        causes <- df %>%
            select(ends_with("_cghr10")) %>%
            pivot_longer(
                everything(),
                names_to = "column",
                values_to = "cod"
            ) %>%
            distinct(cod) %>%
            filter(!is.na(cod)) %>%
            pull(cod)

        # Calc metric by cod
        for (cod in causes) {

            # Filter data for cod
            df_filter <- df %>% filter(physician_cghr10 == cod)
            metric_name = paste0(name, " COD ", cod)

            # Calc metric for cod
            out[[metric_name]] <- calc_per_model(
                actual = df_filter %>%
                    pull(physician_cghr10),
                pred = df_filter %>%
                    select(ends_with("_cghr10"), -physician_cghr10),
                func = func,
                ...
            ) %>% rename(
                 {{ metric_name }} := Metric
            )
        }
    }

    # Combine all pccc metrics
    out <- reduce(out,
        function(x, y) left_join(x, y, by = "Model")
    ) %>%
        arrange(across({{ name }}, desc))
    return(out)
}
```

## Data

### Raw Data

Load data from `data` folder.

```{r}
raw_df <- read_csv("../data/healsl_rd1to2_cod_v1.csv")
```

```{r, echo=FALSE}
cat("Cases: ", nrow(raw_df), "\n")
print(head(raw_df))
```

### Physician Cases

Filter for cases that were coded by physicians.

```{r}
df <- raw_df %>%
    filter(!is.na(physician_cghr10) & !is.null(physician_cghr10))
```

```{r, echo=FALSE}
cat("Physician Coded Cases: ", nrow(df))
print(head(df))
```

### Combined Models

Create model combinations where if any of the models have the physician code, then set the combined model's output for the case to be the physician code.

```{r}
df <- df %>%
    mutate(
        "chatgpt3_insilicova_cghr10" = if_else(
            chatgpt3_cghr10 == physician_cghr10 | 
                insilicova_cghr10 == physician_cghr10,
            physician_cghr10,
            chatgpt3_cghr10
        ),
        "chatgpt4_insilicova_cghr10" = if_else(
            chatgpt4_cghr10 == physician_cghr10 | 
                insilicova_cghr10 == physician_cghr10,
            physician_cghr10,
            chatgpt4_cghr10
        ),
        "chatgpt3_interva5_cghr10" = if_else(
            chatgpt3_cghr10 == physician_cghr10 | 
                interva5_cghr10 == physician_cghr10,
            physician_cghr10,
            chatgpt3_cghr10
        ),
        "chatgpt4_interva5_cghr10" = if_else(
            chatgpt4_cghr10 == physician_cghr10 | 
                interva5_cghr10 == physician_cghr10,
            physician_cghr10,
            chatgpt4_cghr10
        ),
        "chatgpt3_insilicova_interva5_cghr10" = if_else(
            chatgpt3_cghr10 == physician_cghr10 | 
                insilicova_cghr10 == physician_cghr10 |
                interva5_cghr10 == physician_cghr10,
            physician_cghr10,
            chatgpt3_cghr10
        ),
        "chatgpt4_insilicova_interva5_cghr10" = if_else(
            chatgpt4_cghr10 == physician_cghr10 | 
                insilicova_cghr10 == physician_cghr10 |
                interva5_cghr10 == physician_cghr10,
            physician_cghr10,
            chatgpt4_cghr10
        )
    )
```

```{r, echo=FALSE}
print(head(df %>% select(
    chatgpt3_insilicova_cghr10,
    chatgpt4_insilicova_cghr10,
    chatgpt3_interva5_cghr10,
    chatgpt4_interva5_cghr10,
    chatgpt3_insilicova_interva5_cghr10,
    chatgpt4_insilicova_interva5_cghr10
)))
```

### Prepared Data

Display case counts for raw data and prepared data after combining models and filtering for physician coded cases.

```{r}

cat("\nRaw Data\n--------\n\n")

# Physician info
raw_physicians <- raw_df$physician_cghr10
cat(paste0(
    "Cases (",
    length(unique(na.omit(raw_physicians))), " CODs): ",
    length(raw_physicians),
    "\n"
))

# Models cases
raw_models <- raw_df %>% select(ends_with("_cghr10"), -physician_cghr10)
for (mcol in colnames(raw_models)) {
    m <- raw_models[[mcol]]
    cat(paste0(
        str_replace_all(mcol, models_format), " Predicted Cases (", 
        length(unique(na.omit(m))), " CODs): ",
        length(m) - sum(is.na(m) | is.null(m)),
        "\n"
    ))
}

cat("\nPrepared Data\n--------------\n")

# Physician cases
physicians <- df$physician_cghr10
cat(paste0(
    "Physician Coded Cases (",
    length(unique(na.omit(physicians))), " CODs): ",
    length(physicians) - sum(is.na(physicians) | is.null(physicians)),
    "\n"
))

# Models cases
models <- df %>% select(ends_with("_cghr10"), -physician_cghr10)
for (mcol in colnames(models)) {
    m <- models[[mcol]]
    cat(paste0(
        str_replace_all(mcol, models_format), " Predicted Cases (", 
        length(unique(na.omit(m))), " CODs): ",
        length(m) - sum(is.na(m) | is.null(m)),
        "\n"
    ))
}
```

## Methods

Calculate metrics from model outputs compared to physician codes.

```{r}
metrics <- list()
```

### PCCC

Calculate Partial Chance Corrected Concordance (PCCC) to evaluate indivudal performance for each model by age and physician coding stage.

```{r}

# Get num of unique causes
ncauses <- df %>%
    select(ends_with("_cghr10")) %>%
    pivot_longer(
        everything(),
        names_to = "column",
        values_to = "cod"
    ) %>%
    distinct(cod) %>%
    filter(!is.na(cod)) %>%
    pull(cod) %>%
    length

# Calc pccc for age, stage, and cod
metrics$pccc <- calc_by(df, calc_pccc, "PCCC", N = ncauses)
metrics$pccc
```

### CSMF Accuracy

Calculate Cause Specific Mortality Fraction (CSMF) Accuracy to evaluate population performance for each model by age and physician coding stage.

```{r}
metrics$csmf_acc <- calc_by(
    df,
    calc_csmf_acc,
    "CSMF Accuracy",
    by_cod = FALSE
)
metrics$csmf_acc
```

## Results

Save calculated metrics to `data` folder.

```{r}
out <- reduce(metrics,
              function(x, y) left_join(x, y, by = "Model")
)
write_csv(out, "../data/healsl_rd1to2_metrics_v1.csv")
```

### All vs Agree Performance

Display a plot of model performance for all records versus records where physicians agreed on the COD code.

```{r}

# Calc stage counts
nall <- nrow(df)
nagree <- nrow(df %>% filter(is_agreed == TRUE))
nrecon <- nrow(df %>% filter(is_recon == TRUE))
nadj <- nrow(df %>% filter(is_adj == TRUE))

# Calc stage perc
pagree <- (nagree / nall) * 100
precon <- (nrecon / nall) * 100
padj <- (nadj / nall) * 100

# Create stage with counts
stage_remap <- c(
    "All" = paste0("All\nRecords\n(n=", nall, ", 100%)"),
    "Agreed" = paste0("Physician Agreed\nRecords\n(n=", nagree, ", ", round(pagree), "%)"),
    "Reconciled" = paste0("Reconciled\n(n=", nrecon, ", ", round(precon), "%)"),
    "Adjudicated" = paste0("Adjudicated\n(n=", nadj, ", ", round(padj), "%)")
)

# Prepare boxplot data
alldata <- out %>%
    select(
        Model,
        `PCCC`,
        `PCCC Agreement`
    ) %>%
    rename(
         All = `PCCC`,
         Agreed = `PCCC Agreement`
    ) %>%
    filter(!str_detect(Model, "&")) %>% # remove multi models
    pivot_longer( # transform to long format
        cols = -Model,
        names_to = "Stage",
        values_to = "PCCC"
    ) %>%
    mutate( # re-order stage
        Stage = factor(Stage, levels = c(
            "Agreed",
            "All"
        ))
    ) %>%
    mutate( # rename stages with counts
        Stage = recode(Stage, !!!stage_remap)
    ) %>%
    group_by(Stage) %>%
    mutate( # Add min, mid, and max model names to label
        "PCCC Min" = if_else(PCCC <= min(PCCC) * 1.05, paste0(Model, " (", round(PCCC, 2), ")"), NA),
        "PCCC Max" = if_else(PCCC >= max(PCCC) * 0.97, paste0(Model, " (", round(PCCC, 2), ")"), NA),
        "PCCC Mid" = if_else(is.na(`PCCC Min`) & is.na(`PCCC Max`), paste0(Model, "\n(", round(PCCC, 2), ")"), NA),
        "PCCC Min" = if_else(`PCCC` == min(PCCC), paste(na.omit(`PCCC Min`), collapse = "\n"), NA), # Combine into one row if close PCCC
        "PCCC Max" = if_else(`PCCC` == max(PCCC), paste(na.omit(`PCCC Max`), collapse = "\n"), NA)
    )

# Plot boxplot
allplot <- alldata %>%
    ggplot(aes(x = Stage, y = PCCC)) +
    geom_boxplot(
        linewidth = 0.5,
        width = 0.25
    ) +
    geom_dotplot(
        binaxis = "y",
        binwidth = 0.03,
        dotsize = 0.3,
        fill = "white",
        color = "darkgray",
        stackdir = "center",
        stackratio = 2,
        position = position_nudge(x = -0.3)
    ) +
    geom_text(
        aes(label = `PCCC Min`),
        color = "#4d4d4d",
        size = 2,
        hjust = 1,
        position = position_nudge(y = -0.015, x = -0.3)
    ) +
    geom_text(
        aes(label = `PCCC Max`),
        color = "#4d4d4d",
        size = 2,
        position = position_nudge(y = 0.065, x = -0.3)
    ) +
    geom_text(
        aes(label = `PCCC Mid`),
        color = "#4d4d4d",
        size = 2,
        hjust = 1,
        position = position_nudge(y = -0.015, x = -0.3)
    ) +
    stat_summary( # min pccc txt on boxplot
        geom = "text",
        fun = min,
        aes(label = sprintf("%1.2f", after_stat(y))),
        position = position_nudge(y = -0.025),
        size = 3
    ) +
    stat_summary( # max pccc txt on boxplot
        geom = "text",
        fun = max,
        aes(label = sprintf("%1.2f", after_stat(y))),
        position = position_nudge(y = 0.025),
        size = 3
    ) +
    labs(
        x = element_blank(),
        y = "PCCC (0=Low, 1=High)"
    ) +
    ylim(0.2, 0.8) +
    coord_flip() +
    theme_minimal() +
    theme(
        panel.grid = element_blank(),
        axis.line = element_line(color = "black"),
        axis.title.x = element_text(margin = margin(t = 5)),
        axis.title.y = element_text(margin = margin(r = 5)),
        axis.line.x = element_line(
            arrow = grid::arrow(length = unit(0.2, "cm"), ends = "both")
        ),
        axis.line.y = element_blank()
    )

# Save the plot
allplot
ggsave("../manuscript/fig-perf-allvsagree.pdf", plot = allplot, dpi = 300)
```

### COD Performance

Plot model performance for physician agreed records by cause of death category.

```{r, fig.height = 16, fig.width = 8}

# Calc total agreed
nagree <- df %>% filter(is_agreed == TRUE) %>% nrow

# Get unique cause columns
causes_col <- out %>% select(starts_with("PCCC COD ")) %>% names

# Create cod with counts
cod_remap <- list()
for (cod_col in causes_col) {

    # Get cod without prefix
    cod <- gsub("PCCC COD ", "", cod_col)

    # Calculate num and perc cases for cod
    ncod <- df %>% filter(physician_cghr10 == cod & is_agreed == TRUE) %>% nrow
    pcod <- (ncod / nagree) * 100
    pcod_label <- if (round(pcod) >= 1) round(pcod) else "<1"

    # Break cod into newlines if more than 3 words
    if (str_count(cod, "\\s+") > 3) {
        cod_label <- str_replace(cod, "(\\S+\\s+\\S+\\s+\\S+) ", "\\1\n")
    } else {
        cod_label <- cod
    }

    # Add label for cod
    cod_remap[[cod]] <- paste0(cod_label, "\n(n=", ncod, ", ", pcod_label, "%)")
}

# Prepare boxplot data
coddata <- out %>%
    select(
        Model,
        starts_with("PCCC COD ")
    ) %>%
    rename_at(
         vars(-Model),
         ~gsub("PCCC COD ", "", .)
    ) %>%
    filter(!str_detect(Model, "&")) %>% # remove multi models
    pivot_longer( # transform to long format
        cols = -Model,
        names_to = "COD",
        values_to = "PCCC"
    ) %>%
    mutate( # rename cod with counts
        "COD" = recode(`COD`, !!!cod_remap)
    ) %>%
    group_by(`COD`) %>%
    mutate( # Make values less than 0 equal to 0
        PCCC = if_else(PCCC <= 0, 0, PCCC)
    ) %>%
    mutate( # Add min, mid, and max model names to label
        "PCCC Min" = if_else(PCCC <= min(PCCC) * 1.03, paste0(Model, " (", round(PCCC, 2), ")"), NA),
        "PCCC Max" = if_else(PCCC >= max(PCCC) * 0.97, paste0(Model, " (", round(PCCC, 2), ")"), NA),
        "PCCC Mid" = if_else(is.na(`PCCC Min`) & is.na(`PCCC Max`), paste0(Model, "\n(", round(PCCC, 2), ")"), NA),
        "PCCC Min" = if_else(`PCCC` == min(PCCC), paste(na.omit(`PCCC Min`), collapse = "\n"), NA), # Combine into one row if close PCCC
        "PCCC Max" = if_else(`PCCC` == max(PCCC), paste(na.omit(`PCCC Max`), collapse = "\n"), NA)
    )

# Create cod order based on max pccc
codorder <- coddata %>%
    group_by(COD) %>%
    summarise("PCCC Max Value" = max(PCCC, na.rm = TRUE)) %>%
    select(COD, `PCCC Max Value`) %>%
    arrange(desc(`PCCC Max Value`)) %>%
    pull(COD)
coddata$COD <- factor(coddata$COD, levels = rev(codorder))

# Plot boxplot
codplot <- coddata %>%
    ggplot(aes(x = COD, y = PCCC)) +
    geom_boxplot(
        linewidth = 0.5,
        width = 0.25
    ) +
    geom_text(
        aes(label = `PCCC Min`),
        color = "#4d4d4d",
        size = 2.5,
        hjust = 1,
        position = position_nudge(y = -0.12)
    ) +
    geom_text(
        aes(label = `PCCC Max`),
        color = "#4d4d4d",
        size = 2.5,
        hjust = 0,
        position = position_nudge(y = 0.12)
    ) +
    stat_summary( # min pccc txt on boxplot
        geom = "text",
        fun = min,
        aes(label = sprintf("%1.2f", after_stat(y))),
        position = position_nudge(y = -0.06),
        size = 3
    ) +
    stat_summary( # max pccc txt on boxplot
        geom = "text",
        fun = max,
        aes(label = sprintf("%1.2f", after_stat(y))),
        position = position_nudge(y = 0.06),
        size = 3
    ) +
    labs(
        x = paste0("Agreed Records (n=", nagree, ", 100%) for ", length(causes_col), " Causes of Death"),
        y = "PCCC (0=Low, 1=High)"
    ) +
    ylim(-0.3, 1.25) +
    coord_flip() +
    theme_minimal() +
    theme(
        panel.grid = element_blank(),
        axis.line = element_blank(),
        axis.line.y = element_blank(),
        axis.title.x = element_text(margin = margin(t = 5)),
        axis.title.y = element_text(margin = margin(r = 5)),
        axis.line.x = element_line(
            color = "black",
            arrow = grid::arrow(length = unit(0.2, "cm"), ends = "both")
        )
    )

# Save the plot
codplot
ggsave("../manuscript/fig-perf-cod.pdf", plot = codplot, dpi = 300)
```

### Age Group Perfomance

Display a plot of performance for physician agreed records by age group.

```{r}

# Calc age counts
nagree <- nrow(df %>% filter(is_agreed == TRUE))
nadult <- nrow(df %>% filter(age == "adult" & is_agreed == TRUE))
nchild <- nrow(df %>% filter(age == "child" & is_agreed == TRUE))
nneo <- nrow(df %>% filter(age == "neo" & is_agreed == TRUE))

# Calc age perc
padult <- (nadult / nagree) * 100
pchild <- (nchild / nagree) * 100
pneo <- (nneo / nagree) * 100

# Create age with counts
age_remap <- c(
    "Adult" = paste0("Adult\n12+ years\n(n=", nadult, ", ", round(padult), "%)"),
    "Child" = paste0("Child\n28 days to 11 years\n(n=", nchild, ", ", round(pchild), "%)"),
    "Neonatal" = paste0("Neonatal\n<28 days\n(n=", nneo, ", ", round(pneo), "%)")
)

# Prepare boxplot data
agedata <- out %>%
    select(
        Model,
        `PCCC Adult Agreement`,
        `PCCC Child Agreement`,
        `PCCC Neo Agreement`
    ) %>%
    rename(
         Adult = `PCCC Adult Agreement`,
         Child = `PCCC Child Agreement`,
         Neonatal = `PCCC Neo Agreement`
    ) %>%
    filter(!str_detect(Model, "&")) %>% # remove multi models
    pivot_longer( # transform to long format
        cols = -Model,
        names_to = "Age Group",
        values_to = "PCCC"
    ) %>%
    mutate( # re-order age
        "Age Group" = factor(`Age Group`, levels = c(
            "Neonatal",
            "Child",
            "Adult"
        ))
    ) %>%
    mutate( # rename ages with counts
        "Age Group" = recode(`Age Group`, !!!age_remap)
    ) %>%
    group_by(`Age Group`) %>%
    mutate( # Add min, mid, and max model names to label
        "PCCC Min" = if_else(PCCC <= min(PCCC) * 1.15, paste0(Model, " (", round(PCCC, 2), ")"), NA),
        "PCCC Max" = if_else(PCCC >= max(PCCC) * 0.95, paste0(Model, " (", round(PCCC, 2), ")"), NA),
        "PCCC Mid" = if_else(is.na(`PCCC Min`) & is.na(`PCCC Max`), paste0(Model, "\n(", round(PCCC, 2), ")"), NA),
        "PCCC Min" = if_else(`PCCC` == min(PCCC), paste(na.omit(`PCCC Min`), collapse = "\n"), NA), # Combine into one row if close PCCC
        "PCCC Max" = if_else(`PCCC` == max(PCCC), paste(na.omit(`PCCC Max`), collapse = "\n"), NA)
    )

# Plot boxplot
ageplot <- agedata %>%
    ggplot(aes(x = `Age Group`, y = PCCC)) +
    geom_boxplot(
        linewidth = 0.5,
        width = 0.25
    ) +
    geom_dotplot(
        binaxis = "y",
        binwidth = 0.03,
        dotsize = 0.3,
        fill = "white",
        color = "darkgray",
        stackdir = "center",
        stackratio = 2,
        position = position_nudge(x = -0.3)
    ) +
    geom_text(
        aes(label = `PCCC Min`),
        color = "#4d4d4d",
        size = 2,
        hjust = 1,
        position = position_nudge(y = -0.015, x = -0.3)
    ) +
    geom_text(
        aes(label = `PCCC Max`),
        color = "#4d4d4d",
        size = 2,
        position = position_nudge(y = 0.055, x = -0.3)
    ) +
    geom_text(
        aes(label = `PCCC Mid`),
        color = "#4d4d4d",
        size = 2,
        hjust = 1,
        position = position_nudge(y = -0.015, x = -0.3)
    ) +
    stat_summary( # min pccc txt on boxplot
        geom = "text",
        fun = min,
        aes(label = sprintf("%1.2f", after_stat(y))),
        position = position_nudge(y = -0.025),
        size = 3
    ) +
    stat_summary( # max pccc txt on boxplot
        geom = "text",
        fun = max,
        aes(label = sprintf("%1.2f", after_stat(y))),
        position = position_nudge(y = 0.025),
        size = 3
    ) +
    labs(
        x = paste0("Agreed Records (n=", nagree, ", 100%)"),
        y = "PCCC (0=Low, 1=High)"
    ) +
    ylim(0.2, 0.8) +
    coord_flip() +
    theme_minimal() +
    theme(
        panel.grid = element_blank(),
        axis.line = element_line(color = "black"),
        axis.title.x = element_text(margin = margin(t = 5)),
        axis.title.y = element_text(margin = margin(r = 5)),
        axis.line.x = element_line(
            arrow = grid::arrow(length = unit(0.2, "cm"), ends = "both")
        ),
        axis.line.y = element_blank()
    )

# Save the plot
ageplot
ggsave("../manuscript/fig-perf-ages.pdf", plot = ageplot, dpi = 300)
```